---
title: "Text Mining With R"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
params:
  date: !r Sys.Date()      
---

```{r, message = FALSE, warning = FALSE}
library(data.table)
library(tidytext)
library(textdata)
library(syuzhet)
library(stringr)
library(janeaustenr)
library(gutenbergr)
library(grid)
library(gtable)
library(ggplot2)
library(RtextminerPkg)
library(RplotterPkg)
```

```{r,setup, include=FALSE, eval=TRUE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, results="as-is", collapse=TRUE, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8)
```

<div>Author: Rick Dean</div>
<div>Article date: `r params$date`</div>

<div class="abstract">
  <p class="abstract">Abstract</p>
  The following notes/R scripts were inspired from Chapter 2 of [Text Mining With R](https://www.tidytextmining.com/index.html) by Julia Silge and David Robinson
</div>  

<div class="note">Note: Functions from the `data.table` and `RtextminerPkg` packages are used in performing many of the following data set operations.</div>

# Sentiment Analysis with Tidy Data

## The sentiments datasets

### Look at the "afinn" dictionary

```{r}
afinn_dt <- data.table::setDT(textdata::lexicon_afinn())
str(afinn_dt)
```
### Look at the "bing" dictionary
```{r}
bing_dt <- data.table::setDT(tidytext::sentiments)
str(bing_dt)
```
### Look at the  "nrc" dictionary
```{r}
nrc_dt <- data.table::setDT(syuzhet::get_sentiment_dictionary("nrc", language = "english"))
str(nrc_dt)
```

## Sentiment analysis with inner join
### Sentiment analysis in the Jane Austen book `Emma`
1. Get the Austen books:
```{r}
jane_austen_books_dt <- data.table::setDT(janeaustenr::austen_books())
str(jane_austen_books_dt)
```

2. Create a *chapter* column from *jane_austen_books_dt*:
```{r}
jane_austen_books_dt[, `:=`(linenumber = .I, chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE))))]
str(jane_austen_books_dt)
```

3. Using `RtextminerPkg::tokenize_text()` to get the words of Jane Austen books:
```{r}
austen_words_lst <- RtextminerPkg::tokenize_text(
  x = jane_austen_books_dt,
  input_col = "text",
  output_col = "word",
  #stopwords = stopwords::stopwords(language = "en"),
  stopwords = tidytext::stop_words$word,
  strip_punct = TRUE
)
austen_words_lst$tokens_dt[1:10]
```
4. Select the `Emma` words:
```{r}
emma_words_dt <- austen_words_lst$tokens_dt[book == "Emma"]
emma_words_dt[1:10]
```
5. Get the "joy" words from the "nrc" dictionary:
```{r}
joy_words_dt <- nrc_dt[sentiment == "joy"]
joy_words_dt[1:10]
```
6. Inner join the "joy" words with the words in `Emma`:
```{r}
emma_joy_dt <- RtextminerPkg::join_words(emma_words_dt, joy_words_dt)
emma_joy_dt[, .(book, sentiment, value)][1:10]
```
7. Get the frequency of words that are "joy":
```{r}
emma_joy_n_dt <- emma_joy_dt[, .(.N), by = word][order(-N)]
emma_joy_n_dt[1:10]
```

## Sentiment changes

1. Inner join all of the Jane Austen books of words with the `Bing` lexicon:
```{r}
austen_bing_dt <- RtextminerPkg::join_words(austen_words_lst$tokens_dt, bing_dt)
austen_bing_dt[1:10]
```
2. Add an *index* column that reflects each word's location by dividing the *linenumber* variable by integer of 80. The *index* keeps track of which 80-line section of the text we are counting up negative and positive sentiments in:
```{r}
austen_bing_dt[, index := (linenumber %/% 80)]
austen_bing_dt[,.(book, word, sentiment, index)][1:10]
```
3. Group by *book*, *index*, and *sentiment* and get the counts:
```{r}
austen_sentiment_dt <- austen_bing_dt[, .(.N), by = .(book, index, sentiment)]
austen_sentiment_dt[1:10][order(-N)]
```
4. To plot the above results, reshape *austen_sentiment_dt* to a "wide" format where we collect all the *N* values across the same row based on columns of *sentiment*:
```{r}
austen_sentiment_wide_dt <- data.table::dcast(austen_sentiment_dt, book + index ~ sentiment, value.var = "N")
str(austen_sentiment_wide_dt)
```
5. Add a *sentiment_val* column to *austen_sentiment_wide_dt* by taking the difference in the frequency values (i.e. *positive* - *negative*):
```{r}
austen_sentiment_wide_dt[, sentiment_val := (positive - negative)]
austen_sentiment_wide_dt[1:10]
```
6. Plot the *sentiment_val* across the *index* for each of the books:
```{r, fig.height=12, fig.width=13}
RplotterPkg::multi_stick_plot(
  df = austen_sentiment_wide_dt,
  factor_var = "book",
  factor_x = "index",
  aes_y = "sentiment_val",
  title = "Net Difference Between Counts of Positive and Negative Words of Sentiment",
  subtitle = "Net counts from ordered 80 character segments in novels by Jane Austen",
  x_title = "Book Index Location",
  y_titles = rep("Net Sentiment Count",6),
  y_limits = c(-30,30),
  y_major_breaks = seq(-30,30,10),
  rot_y_tic_label = T,
  col_width = 5.5,
  row_height = 3
)
```

<blockquote>The plots show how each novel changes toward more positive or negative sentiment over the trajectory of the story.</blockquote>

## Comparing sentiment lexicons

### Comparing afinn, bing, and nrc lexicons for sentiment classification
<div class="note">Note: See the R script for this topic in the current directory at `compare_lexicons.R`</div>


### Comparing NRC with Bing for the ratio of negative to positive
1. Get the ratio of negetive over positive words in the NRC lexicon:
```{r}
nrc_dt <- data.table::setDT(syuzhet::get_sentiment_dictionary("nrc", language = "english"))
nrc_neg_pos_dt <- nrc_dt[sentiment %in% c("positive", "negative"), .N ,by = sentiment]
str(nrc_neg_pos_dt)
```
2. Take the ratio of negative to positive counts:
```{r}
nrc_neg_pos_ratio <- nrc_neg_pos_dt$N[[2]]/nrc_neg_pos_dt$N[[1]]
nrc_neg_pos_ratio
```

3. Get the ratio of negative over positive words in the Bing lexicon:
```{r}
bing_dt <- data.table::setDT(tidytext::sentiments)
bing_neg_pos_dt <- bing_dt[, .N, by = sentiment]
str(bing_neg_pos_dt)
```
4. Take the ratio of negative to positive counts:
```{r}
bing_neg_pos_ratio <- bing_neg_pos_dt$N[[1]]/bing_neg_pos_dt$N[[2]]
bing_neg_pos_ratio
```

<div class="note">Note: There appears to be higher negative words in both lexicons and with a higher ratio in Bing.</div>

### Most common positive and negative words in a text

1. Find the words across all of Austen's novelsd:
```{r}
jane_austen_books_dt <- data.table::setDT(janeaustenr::austen_books())
austen_words_lst <- RtextminerPkg::tokenize_text(
  x = jane_austen_books_dt,
  input_col = "text",
  output_col = "word"
) 
austen_words_dt <- austen_words_lst$tokens_dt
str(austen_words_dt)
```

2. Do an inner join of all the Austen words with the Bing lexicon:
```{r}
bing_dt <- data.table::setDT(tidytext::sentiments)
austen_bing_dt <- RtextminerPkg::join_words(austen_words_dt, bing_dt)
str(austen_bing_dt)
```
3. Get the top word counts that are common to both the Austen and Bing lexicon words:
```{r}
austen_bing_counts_dt <- austen_bing_dt[, .N, by = .(word, sentiment)][order(-N)]
austen_bing_counts_dt[1:10]
```

4. Plot a bar chart of the top 10 positive and negative words:
```{r, fig.width=12}
austen_bing_positive_dt <- austen_bing_counts_dt[sentiment == "positive"][1:10]
austen_bing_negative_dt <- austen_bing_counts_dt[sentiment == "negative"][1:10]
positive_wrd_plot <- RplotterPkg::create_bar_plot(
  df = austen_bing_positive_dt,
  aes_x = "word",
  aes_y = "N",
  title = "Top Positive Words",
  do_coord_flip = T,
  rot_y_tic_label = T,
  bar_fill = "green",
  order_bars = "asc"
)
negative_wrd_plot <- RplotterPkg::create_bar_plot(
  df = austen_bing_negative_dt,
  aes_x = "word",
  aes_y = "N",
  title = "Top Negative Words",
  do_coord_flip = T,
  rot_y_tic_label = T,
  bar_fill = "red",
  order_bars = "asc"
)
layout <- list(
  plots = list(positive_wrd_plot, negative_wrd_plot),
  rows = c(1,1),
  cols = c(1,2)
)
RplotterPkg::multi_panel_grid(
  layout = layout,
  col_widths = c(5,5),
  row_heights = c(5),
  title = "Word Frequency for Positive and Negative Sentiments",
  subtitle = "Using the Bing lexicon across all Jane Austen novels."
)
```

## Using a regex pattern for tokenization

### Number of chapters in each book
1. Get Jane Auten books:
```{r}
jane_austen_books_dt <- data.table::setDT(janeaustenr::austen_books())
```

2. Tokenize the text using regex to locate rows with a chapter pattern:
```{r}
jane_austen_books_dt <- data.table::setDT(janeaustenr::austen_books())
jane_austen_chapters_dt <- RtextminerPkg::tokenize_text(
  x = jane_austen_books_dt,
  type = "regex",
  input_col = "text",
  output_col = "chapter",
  regex_pattern = "Chapter|CHAPTER [\\dIVXLC]",
  regex_return = "logical"
)

```
3. Sum the number of chapters for each book:
```{r}
jane_auten_chapters_totals_dt <- jane_austen_chapters_dt$tokens_dt[, .(chapters = sum(chapter)), by = book]
print(jane_auten_chapters_totals_dt)
```

### Number of words in each chapter
1. Get the Austen books and add a *chapter* variable:
```{r}
ja_books_dt <- data.table::setDT(janeaustenr::austen_books())
ja_books_dt[, chapter := cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE))), by = book]
str(ja_books_dt)
```
2. Remove the "0" chapter:
```{r}
ja_books_dt <- ja_books_dt[chapter != 0]
str(ja_books_dt)
```

3. Tokenize `ja_books_dt` by word:
```{r}
ja_words_lst <- RtextminerPkg::tokenize_text(
  x = ja_books_dt,
  input_col = "text",
  output_col = "word"
) 
ja_words_dt <- ja_words_lst$tokens_dt
str(ja_words_dt)
```

4. Get the number of words in each chapter:
```{r}
ja_chapter_wrds_n_dt <- ja_words_dt[, .(words = .N), by = .(book, chapter)]
str(ja_chapter_wrds_n_dt)
```
### Get the number of negative Bing words in each chapter
1. Get the negative words from Bing lexicon:
```{r}
bing_dt <- data.table::setDT(tidytext::sentiments)
bing_neg_dt <- bing_dt[sentiment == "negative"]
str(bing_neg_dt)
```

2. Inner join the Austen words with the negative Bing words:
```{r}
ja_neg_dt <- RtextminerPkg::join_words(ja_words_dt, bing_neg_dt)
str(ja_neg_dt)
```
3. Find the number of negative words in each chapter:
```{r}
ja_neg_chap_n_dt <- ja_neg_dt[, .(neg_words = .N), by = .(book, chapter)]
str(ja_neg_chap_n_dt)
```
4. Outer join the number of words in each chapter (*ja_chapter_wrds_n_dt*) with the number of negative words in each chapter (*ja_neg_chap_n_dt*):
```{r}
data.table::setkey(ja_chapter_wrds_n_dt, book, chapter)
data.table::setkey(ja_neg_chap_n_dt, book, chapter)
ja_neg_join_dt <- ja_chapter_wrds_n_dt[ja_neg_chap_n_dt, nomatch = 0]
str(ja_neg_join_dt)
```
5. Add a *ratio* variable dividing *neg_words* by *words*:
```{r}
ja_neg_join_dt[, ratio := neg_words/words][order(-ratio)][1:10]
```
6. Show Emma, chapter 15:
```{r}
ja_neg_join_dt[book == "Emma" & chapter == 15]
```

