---
title: "Text Mining With R"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
params:
  date: !r Sys.Date()      
---

```{r, message = FALSE, warning = FALSE}
library(data.table)
library(janeaustenr)
library(tidytext)
library(gutenbergr)
library(stringr)
library(grid)
library(gtable)
library(rlang)
library(scales)
library(ggplot2)
library(ggrepel)
library(RtextminerPkg)
library(RplotterPkg)
```

```{r,setup, include=FALSE, eval=TRUE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, results="as-is", collapse=TRUE, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8)
```

<div>Author: Rick Dean</div>
<div>Article date: `r params$date`</div>

<div class="abstract">
  <p class="abstract">Abstract</p>
  The following notes/R scripts were inspired from Chapter 1 of [Text Mining With R](https://www.tidytextmining.com/index.html) by Julia Silge and David Robinson
</div>  

<div class="note">Note: Functions from the `data.table` and `RtextminerPkg` packages are used in performing many of the following data set operations.</div>

# The Tidy Text Format

## Tokenize text into words

### Words from Emily Dickinson text
1. The data:
```{r}
emily_vec <- c("Because I could not stop for Death",
          "He kindly stopped for me",
          "The Carriage held but just Ourselves",
          "and Immortality")
emily_df <- data.frame(
  line = 1:4,
  text = emily_vec
)
print(emily_df)
```
2. Using `RtextminerPkg::tokenize_text()`:
```{r}
emily_words_lst <- RtextminerPkg::tokenize_text(
  x = emily_df,
  type = "word",
  output_col = "word"
)
```
The returned list *emily_words_lst* is a named list with: 

1. A `data.table::data.table` named *tokens_dt* with columns for the text's line number and corresponding tokenized words from *emily_df*:
```{r}
emily_words_lst$tokens_dt[1:10]
```

2. A `data.table::data.table` named *tokens_count* showing the frequency of each word in *emily_df*:
```{r}
emily_words_lst$tokens_count[1:6]
```

### Words from the works of Jane Austen
1. The data:
```{r}
jane_austen_books_dt <- data.table::setDT(janeaustenr::austen_books())
str(jane_austen_books_dt)
```
<div class="note">Note: *jane_austen_books_dt* is a data.table and data.frame with a column of character vectors *text* where each vector has a length of 1. </div>

2. Add a *linenumber* and *chapter* columns to *jane_austen_books_dt*:
```{r}
jane_austen_books_dt[, `:=`(linenumber = .I, chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE))))]
str(jane_austen_books_dt)
```

3. Using `RtextminerPkg::tokenize_text()` to get the words of Jane Austen books:
```{r}
austen_words_lst <- RtextminerPkg::tokenize_text(
  x = jane_austen_books_dt,
  input_col = "text",
  output_col = "word",
  stopwords = tidytext::stop_words$word,
  strip_punct = TRUE
)
```
4. Show the words from the text:
```{r}
austen_words_lst$tokens_dt[1:10]
```
5. Show the counts of the words across all the Austen books:
```{r}
austen_words_lst$tokens_count[1:6]
```

6. Show the word counts in a bar chart:
```{r}
RplotterPkg::create_bar_plot(
  df = austen_words_lst$tokens_count[1:13],
  aes_x = "word",
  aes_y = "N",
  title = "Words Counts Across All of Jane Austen's Books",
  x_title = "Word",
  rot_y_tic_label = T,
  do_coord_flip = T,
  bar_labels = T,
  order_bars = "asc",
  bar_fill = "blue",
  bar_color = "gold",
  bar_alpha = 0.6,
  bar_width = 0.7
)
```

### Words from the works of H.G. Wells using `gutenbergr`package

1. Download the books using `gutenbergr`:
```{r}
hgwells_books_dt <- data.table::setDT(gutenbergr::gutenberg_download(c(35,36,5230,159)))
str(hgwells_books_dt)
```
2. Using `RtextminerPkg::tokenize_text()` to get the words of H. G. Wells:
```{r}
hgwells_words_lst <- RtextminerPkg::tokenize_text(
  x = hgwells_books_dt,
  input_col = "text",
  output_col = "word",
  #stopwords = stopwords::stopwords(language = "en"),
  stopwords = tidytext::stop_words$word,
  strip_punct = TRUE
)
```
3. Show the counts of the words across all the H.G. Wells books:
```{r}
hgwells_words_lst$tokens_count[1:6]
```
### Words from the works of Bronte sisters using `gutenbergr`package
1. Download the books using `gutenbergr`:
```{r}
bronte_books_dt <- data.table::setDT(gutenbergr::gutenberg_download(c(1260, 768, 969, 9182, 767)))
str(bronte_books_dt)
```
2. Using `RtextminerPkg::tokenize_text()` to get the words of Bronte Sisters:
```{r}
bronte_words_lst <- RtextminerPkg::tokenize_text(
  x = bronte_books_dt,
  input_col = "text",
  output_col = "word",
  #stopwords = stopwords::stopwords(language = "en"),
  stopwords = tidytext::stop_words$word,
  strip_punct = TRUE
)
```
3. Show the counts of the words across all the Bronte sisters books:
```{r}
bronte_words_lst$tokens_count[1:6]
```
## Word counts by author/word grouping
1. Create data.tables for each author by selecting the *word* column from the above author's *tokens_dt* data.table and creating an *author* column:
```{r}
austen_words_dt <- austen_words_lst$tokens_dt[, .(word = word, author = "JaneAusten")]

hgwells_words_dt <- hgwells_words_lst$tokens_dt[, .(word = word, author = "HGWells")]

bronte_words_dt <- bronte_words_lst$tokens_dt[, .(word = word, author = "BronteSisters")]
```
2. Row bind all of the above word data tables:
```{r}
words_combined_dt <- data.table::rbindlist(list(
  austen_words_dt, 
  hgwells_words_dt, 
  bronte_words_dt))
str(words_combined_dt)
```
3. As a check group *words_combined_dt* by (*author*, *word*) and get the *word* counts:
```{r}
author_word_counts_dt <- words_combined_dt[, .(.N), by = .(author,word)][order(-N)]
author_word_counts_dt[1:10]
```
<div class="note">Note: Compare these *N*'s with the *N*'s we obtained earlier above.</div>

## Word proportions by author/word grouping
1. Group *author_word_counts_dt* by author and divide *N* by the sum of *N*:
```{r}
author_word_proportions_dt <- author_word_counts_dt[, .(word, N, prop = N/sum(N)), by = .(author)][order(-prop)]
author_word_proportions_dt[1:10]
```
## Create a scatter plot of word proportions between authors
1. Turn *author_word_proportions_dt* data table into a "wide" data table where we collect all the *prop* values across the same row based on columns of *author*.
```{r}
wide_author_dt <- data.table::dcast(author_word_proportions_dt, word ~ author, value.var = "prop")
str(wide_author_dt)
```
2. As a check order the proportions of *JaneAusten* in *wide_author_dt*:
```{r}
wide_author_dt[order(-JaneAusten)][1:10]
```
3. Filter *wide_author_dt* to proportions greater than 0.001:
```{r}
wide_author_filter_dt <- wide_author_dt[JaneAusten > 0.001 & BronteSisters > 0.001 & HGWells > 0.001]
wide_author_filter_dt[1:10]
```
4. Plot the proportions from *wide_author_filter_dt* in scatter plots of *JaneAusten* vs *BronteSisters* and *JaneAusten* vs *HGWells*:
```{r}
RplotterPkg::multi_scatter_plot(
  df = wide_author_filter_dt,
  title = "Jane Austen vs Bronte Sisters and H.G. Wells",
  subtitle = "Proportions of words from all their books",
  variables = c("BronteSisters", "HGWells"),
  aes_y = "JaneAusten",
  aes_label = "word",
  x_log10 = TRUE,
  y_log10 = TRUE,
  x_labels = scales::percent_format(),
  y_labels = scales::percent_format(),
  rot_y_tic_label = T,
  pts_fill = "green",
  pts_size = 2,
  col_width = 5.5,
  row_height = 5.5
)
```
